---
title: "Getting and Cleaning Data"
output: html_notebook
---
#Download the file to load
```{r}
# check for file exist or not
if(!file.exists("data")){
  dir.create("data")
}
# download file
fileURL <- "https://data.iowa.gov/api/views/4bfg-n52x/rows.csv?accessType=DOWNLOAD"
download.file(fileURL, destfile = "cameras.csv", method = "curl")
#Check date
dateDownloaded <- date()
```

# mySQL
- Free and widely used open source database software
- Widely used in internet based applications
- Data are structured in 
  - Databases
  - Tables within databases
  - Fields within tables
- Each row called a record
More in : http://en.wikipedia.org/wiki/MySQL or http://www.mysql.com/
## Connecting and listing databases
Install 2 packages:
```{r}
library(RMySQL)
library(DBI)
```
Connect to web have SQL databases:
```{r}
ucscDb <- dbConnect(MySQL(), user = "genome",
                 host ="genome-mysql.cse.ucsc.edu")
result <- dbGetQuery(ucscDb, "show databases;"); dbDisconnect(ucscDb);
```
List databases:
```{r}
result
```
## Connecting to hg19 and listing tables
```{r}
hg19 <- dbConnect(MySQL(), user = "genome", db = "hg19", host = "genome-mysql.cse.ucsc.edu")
allTables <- dbListTables(hg19)
length(allTables)
```
listing tables:
```{r}
allTables[1:5]
```

## Get dimensions of a a specific table
```{r}
dbListFields(hg19,"affyU133Plus2")
dbGetQuery(hg19, "select count(*) from affyU133Plus2")
```
## Read from the table
```{r}
affyData <- dbReadTable(hg19, "affyU133Plus2")
head(affyData)
```
## Select a specific subset
```{r}
query <- dbSendQuery(hg19, "select * from affyU133Plus2 where misMatches between 1 and 3")
affyMis <- fetch(query); quantile(affyMis$misMatches)

affyMissSmall <- fetch(query, n = 10); dbClearResult(query); # Clear query before add new query
dim(affyMissSmall)
```

