---
title: "Reading from HDF5"
output: html_document
date: "2025-01-28"
author:"TRAN ANH MINH"
---

#Introduction

HDF5 (Hierarchical Data Format version 5) is an advanced file format widely used in data analysis and scientific computing:

- Specifically designed for storing and managing large datasets (big data)
- Supports diverse data types: integers, floating-point numbers, strings, etc.
- Flexible hierarchical structure, similar to a file system
- Optimized read/write performance
- Built-in data compression capabilities

## HDF5 Structure:
1. Groups:
   - Function like directories, containing datasets and metadata
   - Group Header: Stores group name and attributes
   - Group Symbol Table: Manages list of objects

2. Datasets:
   - Store data as multidimensional arrays
   - Dataset Header: Defines name, datatype, dataspace
   - Data Array: Contains actual data


# RHDF5 Setup and Initialization
```{r}
# Install BiocManager if not present
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
# Install rhdf5 package
BiocManager::install("rhdf5")
# Load library
library(rhdf5)
# Create new HDF5 file
created = h5createFile("example.h5")
created 
```
# Creating Group Structure
```{r}
# Create main groups 'foo' and 'baa'
created_1 <- h5createGroup("example.h5", "foo")
created_1 <- h5createGroup("example.h5", "baa")

# Create subgroup 'foobaa' in 'foo'
created_1 <- h5createGroup("example.h5", "foo/foobaa")

# Examine file structure
h5ls("example.h5")
```
#Working with Datasets
```{r}
# Create and write 5x2 matrix
A = matrix(1:10, nr=5, nc=2)
h5write(A, "example.h5", "foo/A")

# Create and write 3D array with attributes
B = array(seq(0.1,2.0,by=0.1), dim=c(5, 2, 2))
attr(B, "scale") <- "liter"  # Add metadata
h5write(B, "example.h5","foo/foobaa/B")

# Check structure after adding data
h5ls("example.h5")
```
#Handling Data Frames
```{r}
# Create data frame with multiple data types
df <- data.frame(
    ID = 1L:5L,  # Integer
    Value = seq(0,1, length.out=5),  # Numeric
    Label = c("ab","cde", "ghi", "a", "s"),  # Character
    stringsAsFactors = FALSE
)
h5write(df, "example.h5","df")
h5ls("example.h5")
```
#Reading from HDF5
```{r}
# Read datasets from different groups
readA <- h5read("example.h5","foo/A")
readB <- h5read("example.h5", "foo/foobaa/B")
readA
```
# Partial Data Updates (Partial I/O)
```{r}
# Update first 3 elements of column 1 in matrix A
h5write(c(12, 13, 14),"example.h5", "foo/A", index=list(1:3,1))

# Read and verify results
h5read("example.h5","foo/A")
```





