---
title: "Getting and Cleaning Data in R: Working with MySQL Databases"
author: "TRAN ANH MINH"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_float: true
    theme: cosmo
---
# SECTION 1: DATA ACQUISITION
This section handles the initial data download and setup procedures
Note: The data is downloaded from Iowa government's open data portal
## Create data directory if it doesn't exist
```{r}

if(!file.exists("data")){
  dir.create("data")
  print("Created new 'data' directory")
}
```
## Download and save the dataset
```{r}

fileURL <- "https://data.iowa.gov/api/views/4bfg-n52x/rows.csv?accessType=DOWNLOAD"
download.file(fileURL, destfile = "cameras.csv", method = "curl")
```
## Record download timestamp for reproducibility
```{r}

dateDownloaded <- date()
print(paste("Data downloaded on:", dateDownloaded))
```

# SECTION 2: MySQL DATABASE INTERACTION
This section demonstrates interaction with MySQL databases focusing on the UCSC Genome Database as an example
## 2.1 Required Libraries ----
Install these packages if not already installed:

Install 2 packages:
```{r}
#install.packages(c("RMySQL", "DBI"))
library(RMySQL)
library(DBI)
```
## 2.2 Database Connection and Overview ----
MySQL is:
- A free, open-source database management system
- Widely used in web applications and bioinformatics
- Structured hierarchically: Databases > Tables > Fields
- Each row represents a record
 
Reference: 
- MySQL documentation: http://www.mysql.com/
- Wikipedia: http://en.wikipedia.org/wiki/MySQL



### Establish connection to UCSC Genome Database
```{r}
ucscDb <- dbConnect(MySQL(), 
                   user = "genome",
                   host = "genome-mysql.cse.ucsc.edu")
# Query available databases
result <- dbGetQuery(ucscDb, "show databases;")
dbDisconnect(ucscDb)  # Good practice to disconnect when done
```


### Display available databases
```{r}
print("Available Databases:")
result
```
## 2.3 Working with the hg19 Database ----
### Connect to specific database (hg19)
```{r}
hg19 <- dbConnect(MySQL(), 
                 user = "genome", 
                 db = "hg19", 
                 host = "genome-mysql.cse.ucsc.edu")
allTables <- dbListTables(hg19)
length(allTables)
```
### Get overview of tables
```{r}
allTables <- dbListTables(hg19)
print(paste("Total number of tables:", length(allTables)))
print("First 5 tables:")
allTables[1:5]
```

## 2.4 Exploring the affyU133Plus2 Table ----
### List fields in the table
```{r}
print("Fields in affyU133Plus2:")
dbListFields(hg19,"affyU133Plus2")
```
### Get table dimensions
```{r}
print("Total records in affyU133Plus2:")
dbGetQuery(hg19, "select count(*) from affyU133Plus2")
```
### Preview table data
```{r}
print("Preview of affyU133Plus2 data:")
affyData <- dbReadTable(hg19, "affyU133Plus2")
head(affyData)
```
## 2.5 Advanced Queries ----
### Query records with specific mismatch criteria
```{r}
query <- dbSendQuery(hg19, 
                    "select * from affyU133Plus2 where misMatches between 1 and 3")

```
### Analyze results
```{r}
affyMis <- fetch(query)
print("Quantiles of misMatches:")
quantile(affyMis$misMatches)
```
### Fetch smaller subset for detailed examination
```{r}
affyMissSmall <- fetch(query, n = 10)
dbClearResult(query)  # Clean up query
print("Dimensions of small subset:")
dim(affyMissSmall)
```
### Close database connection
```{r}
dbDisconnect(hg19)
```


